%$Id: r.tex,v 20.14 2001/10/19 06:21:44 al Exp $
% man circuit r .
% Copyright (C) 2001 Albert Davis
% Author: Albert Davis <aldavis@ieee.org>
%
% This file is part of "GnuCap", the Gnu Circuit Analysis Package
%
% This program is free software; you can redistribute it and/or modify
% it under the terms of the GNU General Public License as published by
% the Free Software Foundation; either version 2, or (at your option)
% any later version.
%
% This program is distributed in the hope that it will be useful,
% but WITHOUT ANY WARRANTY; without even the implied warranty of
% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
% GNU General Public License for more details.
%
% You should have received a copy of the GNU General Public License
% along with this program; if not, write to the Free Software
% Foundation, Inc., 59 Temple Place - Suite 330, Boston, MA
% 02111-1307, USA.
%------------------------------------------------------------------------
\section{{\tt R}: Resistor}
%------------------------------------------------------------------------
\subsection{Syntax}
\begin{verse}
  {\tt R}{\it xxxxxxx n+ n-- value}\\
  {\tt R}{\it xxxxxxx n+ n-- expression}\\
  {\tt R}{\it xxxxxxx n+ n-- model} \{{\tt L=}{\it length}\}
        \{{\tt W=}{\it width}\} \{{\tt TEMP=}{\it temperature}\} \\
  {\tt .RESistor} {\it label n+ n-- expression}
\end{verse}
%------------------------------------------------------------------------
\subsection{Purpose}

Resistor, or general current controlled dissipative element.
%------------------------------------------------------------------------
\subsection{Comments}

{\it N+} and {\it n--} are the positive and negative element nodes,
respectively.  {\it Value} is the resistance in Ohms.

The resistor (type {\tt R}) differs from the admittance (type {\tt Y}) in
that the resistor is a current controlled element, and the conductance is a
voltage controlled element, in addition to the obvious use of conductance
($1/R$) instead of resistance.

You may specify the {\it value} in one of three forms:

\begin{enumerate}
  
\item A simple value.  This is the resistance in Ohms.
  
\item An expression, as described in the behavioral modeling chapter.
  The expression can specify the voltage as a function of current, or
  the resistance as a function of time.
  
\item A {\it model}, which calculates the resistance as a function of
  length and width, referencing a {\tt .model} statement of type {\tt
    R}.  This is compatible with the Spice-3 ``semiconductor
  resistor''.

\end{enumerate}
%------------------------------------------------------------------------
\subsection{Model statement}

A model statement may be used,, with model type {\tt R} or {\tt Res}.
The parameters are:

\begin{description}
  
\item[{\tt RSH} = {\it x}] Sheet resistance. (Ohms / square). (Required)
  
\item[{\tt CJSW} = {\it x}] Junction sidewall capacitance. (Farads /
  meter).  (Default = 0.)
  
\item[{\tt DEFW} = {\it x}] Default width. (meters).  (Default = 1e-6)
  
\item[{\tt NARROW} = {\it x}] Narrowing due to side etching. (meters).
  (Default = 0.)
  
\item[{\tt TC1} = {\it x}] First order temperature coefficient.
  (Farads / degree C).  (Default = 0.)
  
\item[{\tt TC2} = {\it x}] Second order temperature coefficient.
  (Farads / degree C squared).  (Default = 0.)
  
\item[{\tt TNOM} = {\it x}] Parameter measurement temperature.
  (degrees C.).  (Default = 27.)

\end{description}

Resistance is computed by the formula:

\begin{verbatim}
resistance = RSH * (L - NARROW) / (W - NARROW)
\end{verbatim}

After the nominal value is calculated, it is adjusted for temperature
by the formula:

\begin{verbatim}
value *= (1 + TC1 * (T-T0) + TC2 * (T-T0)^2)
\end{verbatim}
%------------------------------------------------------------------------
%------------------------------------------------------------------------
