%$Id: vccap.tex,v 25.95 2006/08/26 01:26:53 al Exp $ -*- LaTeX -*-
% man circuit g .
% Copyright (C) 2001 Albert Davis
% Author: Albert Davis <aldavis@ieee.org>
%
% This file is part of "Gnucap", the Gnu Circuit Analysis Package
%
% This program is free software; you can redistribute it and/or modify
% it under the terms of the GNU General Public License as published by
% the Free Software Foundation; either version 2, or (at your option)
% any later version.
%
% This program is distributed in the hope that it will be useful,
% but WITHOUT ANY WARRANTY; without even the implied warranty of
% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
% GNU General Public License for more details.
%
% You should have received a copy of the GNU General Public License
% along with this program; if not, write to the Free Software
% Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
% 02110-1301, USA.
%------------------------------------------------------------------------
\section{Voltage Controlled Capacitor}
%------------------------------------------------------------------------
\subsection{Syntax}
\subsubsection{Device}
\begin{verse}
{\tt G}{\it xxxxxxx n+ n--} VCCAP {\it nc+ nc--} {\it expression}\\
{\tt .VCCAP} {\it label n+ n-- nc+ nc--} {\it expression}
\end{verse}
\subsubsection{Model (optional)}
\begin{verse}
{\tt .model} {\it mname} {\tt C} \{{\it args}\} \\
{\tt .model} {\it mname} {\tt TABLE} \{{\it args}\}
\end{verse}
%------------------------------------------------------------------------
\subsection{Purpose}

Voltage controlled capacitor.
%------------------------------------------------------------------------
\subsection{Comments}

{\it N+} and {\it n--} are the positive and negative element (output) nodes,
respectively.  {\it Nc+} and {\it nc--} are the positive and negative
controlling nodes, respectively.  {\it Value} is the transfactor in
Farads per volt.

The simulator will faithfully give you a negative capacitor if it
seems appropriate.  Usually, this part is used with a behavioral
modeling function, like PWL, which allows you to specify a table of
capacitance vs. voltage.

You may specify the {\it value} in any of these forms:

\begin{enumerate}
  
\item
A simple value.  This is the capacitance in Farads with 1 volt applied
to the control port.
  
\item
An expression, as described in the behavioral modeling chapter.

\item
A {\it model}, as described in the behavioral modeling chapter.  The
{\tt table} model describes a table of capacitance vs. control voltage.

\end{enumerate}
%------------------------------------------------------------------------
\subsection{Probes}

All probes that apply to ordinary capacitors also apply here.
%------------------------------------------------------------------------
%------------------------------------------------------------------------
